### 2021-05-07 manuscript.Rmd
icc analysis added to pipeline

---
title: "Predicting prolonged apnea during nurse-administered procedural sedation using machine learning"
bibliography: "Extras/references.bib"
csl: "Extras/jmir.csl"
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Katina Corones PhD:
      institute: [Griffith]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false

institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - Griffith: School of Nursing, Giffith University, Brisbane, Australia

output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE, warning=FALSE, fig.retina=8)
```

## Abstract

### Background

### Objective

### Methods

### Results

### Conclusion

**Keywords**

Procedural sedation and analgesia; conscious sedation; nursing; patient safety; anesthesia

\pagebreak

## Introduction